---
title: "MatrixCalculation.Rmd"
author: "April Wright"
date: "2023-12-18"
output: html_document
---

We're going to work through comparing the simulated Mk matrices to the empirical, then the simulated punc to the empirical punc. First, I will walk through the reading of data and the general functions so that this is clear, then we'll vectorize the code to operate across a fuckton of simulations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
library(Revticulate)
knitRev()
```

First, read in a matrix. I'm doing this in Revticulate because RevBayes makes concatenation easier than R. 

```{rb}
data <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[1].nex")
data1 <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[2].nex")
data2 <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[3].nex")

data_final <- concatenate(data, data1)
data_final <- concatenate(data_final, data2)
writeNexus(data_final, "test")
```

Now we'll load it in R:

```{r}
data <- ape::read.nexus.data("test")
```

We want to look at five different metrics for evaluating the simularity of matrices:

+ Simple pairwise disparity
+ Gower's 
+ Consistency
+ Retention
+ MP score

Down the row we go.

## Simple Pairwise Disparity

```{r}
Claddis::calculate_morphological_distances(data, distance_metric = "red")

```

## Gower's 
```{r}
Claddis::calculate_morphological_distances(data, distance_metric = "gc")
```

## Consistency and Retention Indices

```{r}
phydata <- read.phyDat("test", format = "nexus", type = "STANDARD")
mcc <- read.nexus("Mk_output/mcc.tre")
phangorn::CI(data=phydata, mcc)
phangorn::RI(data=phydata, mcc)
```

## MP Score

```{r}
phangorn::fitch(tree = mcc, data = phydata, site = "pscore")
#or
#phangorn::sankoff(tree = mcc, data = phydata, site = "pscore")
```


