---
title: "MatrixCalculation.Rmd"
author: "April Wright"
date: "2023-12-18"
output: html_document
---

We're going to work through comparing the simulated Mk matrices to the empirical, then the simulated punc to the empirical punc. First, I will walk through the reading of data and the general functions so that this is clear, then we'll vectorize the code to operate across a fuckton of simulations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
library(Revticulate)
knitRev()
```

First, read in a matrix. I'm doing this in Revticulate because RevBayes makes concatenation easier than R. We read in the matrix with the most states first so the state space initializes correctly (0-1-2-3).

```{rb}
data <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[3].nex")
data1 <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[2].nex")
data2 <- readDiscreteCharacterData("PPS_output/posterior_predictive_sim_1/phyMorpho[1].nex")
data_final <- concatenate(data, data1)
data_final <- concatenate(data_final, data2)
writeNexus(data_final, "test")
```
We want to look at five different metrics for evaluating the simularity of matrices:

+ Simple pairwise disparity
+ Gower's 
+ Consistency
+ Retention
+ MP score

Down the row we go.

## Simple Pairwise Disparity

Claddis is funny about format, so we'll need to do a few things to make the matrix work with it.

```{r}

data_manufacture <- function(data){
  ass_block <- "
  BEGIN ASSUMPTIONS;
	  OPTIONS  DEFTYPE=unord PolyTcount=MINSTEPS ;
	  TYPESET * UNTITLED  = unord: 1-63;
  END;"

  data <- read_file(data)
  data <- toupper(data)
  full <- capture.output(cat(data, ass_block))
  write(full,  "new_test.nex")
  claddis_formatted_data <- Claddis::read_nexus_matrix("new_test.nex")
return(claddis_formatted_data)
  }

```

```{r}
simple <- function(claddis_formatted_data){
  simple_parwise <- Claddis::calculate_morphological_distances(claddis_formatted_data, distance_metric = "red")

  simple_df <- as.data.frame(simple_parwise)
  simple_summary <- simple_df %>% 
    summarise_all(mean)
  mean_simple <- mean(unlist(simple_summary[-1]))
  return(mean_simple)
}
```

## Gower's 
```{r}
gower <- function(claddis_formatted_data){

  gowers <- Claddis::calculate_morphological_distances(claddis_formatted_data, distance_metric = "gc")

  gowersdf <- as.data.frame(gowers)
  gowers_summary <- gowersdf %>% 
    summarise_all(mean)

  mean_gowers <- mean(unlist(gowers_summary[-1]))
}
```

## Consistency and Retention Indices

```{r}
ci <- function(phyData, mcc){
  ci <- phangorn::CI(data=phydata, mcc)
  return(ci)
}
```

```{r}
ri <- function(phyData, mcc){
  ri <- phangorn::RI(data=phydata, mcc)
  return(ri)

}
```

## MP Score

```{r}
mp <- function(phyData, mcc){
  pars <- phangorn::fitch(tree = mcc, data = phydata, site = "pscore")
  return(pars)
}

```

## Automation across multiple sets

Let's allocate vectors for our output
```{r}
simple_pairwise <- c()
gowers <- c()
const <- c()
ret <- c()
mpars_list <- c()
```

Now, we do all the shit, append the results to vector and make a dataframe.

```{r include=FALSE}
dir_list <- list.dirs("PPS_output/")
dir_list <- dir_list[ grepl("post*", dir_list) ]
dir_list <- dir_list[2:102]
for (x in dir_list){
  spl <- strsplit(x, "_")[[1]][5]
  print(spl)
  file1 <- paste0(x, "/phyMorpho[3].nex")
  file2 <- paste0(x, "/phyMorpho[2].nex")
  file3 <- paste0(x, "/phyMorpho[1].nex")
  file.copy(from = file1,
  to   = ".")
  file.copy(from = file2,
  to   = ".")
  file.copy(from = file3,
  to   = ".")
  nex_file <- paste0("test", spl, ".nex")
  system("~/software/rb RevFormat.Rev")
  file.copy(from = "test", to = nex_file)
  data_manufacture(nex_file)
  sdis <- suppressWarnings(simple(claddis_formatted_data))
  simple_pairwise <- c(simple_pairwise, sdis)
  gos_dis <- suppressWarnings(gower(claddis_formatted_data))
  gowers <- c(gowers, gos_dis)
  phydata <- read.phyDat(nex_file, format = "nexus", type = "STANDARD")
  mcc <- read.nexus("Mk_output/mcc.tre")
  cons <- ci(phyData, mcc)
  const <- c(const, cons)
  rent <- ri(phyData, mcc)
  ret <- c(rent, ret)
  mp_Score <- mp(phyData, mcc)
  mpars_list <- c(mpars_list, mp_Score)
}
```

Now, populate a data frame with all of these values, and export it.

```{r}
df <- cbind(simple_pairwise, gowers, const, ret, mpars_list)
write.csv(df, "pps_mk.csv")
```
